---
title: "Intro to Grammar of Graphics (ggplot2)"
author: "Jason Bryer"
date: "September 9, 2014"
output:
  beamer_presentation:
    keep_tex: true
    fig_width: 7
    fig_height: 4
    fig_caption: false
---


## ggplot2

`ggplot2` (Wickham, 2009) is an R package that implements the *Graammar of Graphics* introduced by Wilkinson (2005).

(Journal of Statistical Software paper)[http://www.jstatsoft.org/v17/b03/paper]

## The Lego Package

```{r options, echo=FALSE, results='hide'}
options(width=70)
```

```{r install, eval=FALSE}
# https://github.com/seankross/lego
devtools::install_github("seankross/lego")
```
```{r setup, results='hide', message=FALSE}
library(lego)
library(psych)
```

## The Lego Package

```{r}
data(legosets)
head(legosets, n=3)
```

## Parts of a ggplot2 statement

> - Data  
  ggplot(myDataFrame, aes(x=x, y=y)
> - Layers  
  geom_point(), geom_histogram()
> - Facets  
  facet_wrap(~ cut), facet_grid(~ cut)
> - Scales  
  scale_y_log10()
> - Other options  
  ggtitle('my title'), ylim(c(0, 10000)), xlab('x-axis label')

## There are Lots of Geoms

```{r geoms}
ls("package:ggplot2")[substr(ls("package:ggplot2"), 
							 1, 5) == 'geom_']
```


## Histograms

```{r, fig.height=4, warning=FALSE}
ggplot(legosets, aes(x=Pieces)) + 
	geom_histogram()
```

## Log Transformations

```{r, warning=FALSE}
ggplot(legosets, aes(x=Pieces)) + 
	geom_histogram() + scale_x_log10()
```

## Barplots

```{r, warning=FALSE}
ggplot(legosets, aes(x=Theme)) + geom_bar() + 
	theme(axis.text.x=element_text(angle=90))
```

## Barplots Flipping Coordinates

```{r, warning=FALSE}
ggplot(legosets, aes(x=Theme)) + geom_bar() + 
	coord_flip()
```

## Labeling Barplots

```{r, warning=FALSE}
df <- as.data.frame(table(legosets$Theme))
df <- df[df$Freq > 20,]
ggplot(df, aes(x=Var1, y=Freq, label=Freq)) + 
	geom_bar(stat='identity', alpha=.5) + 
	coord_flip() + geom_text(hjust=0)
```

## Boxplots

```{r, warning=FALSE}
ggplot(legosets, aes(x=Availability, Pieces)) + 
	geom_boxplot()
```

## Boxplots Flipping Coordinates

```{r, warning=FALSE}
ggplot(legosets, aes(x=Availability, Pieces)) + 
	geom_boxplot() + coord_flip()
```

## Boxplots for Longitudinal Data

```{r, warning=FALSE}
ggplot(legosets, aes(x=factor(Year), y=USD_MSRP)) + 
	geom_boxplot() + theme(axis.text.x=element_text(angle=90)) + 
	xlab('Year') + ylab('Price')
```

## Scatterplots

```{r, warning=FALSE}
ggplot(legosets, aes(x=Pieces, y=USD_MSRP)) + 
	geom_point(alhpa=.5)
```

## Scatterplots with Loess Plots

```{r, warning=FALSE}
ggplot(legosets, aes(x=Pieces, y=USD_MSRP)) + 
	geom_point(alhpa=.5) + geom_smooth()
```

